---
title: "GSA 2022 Abstract Code"
author: "Peter Sun"
date: "March 3, 2022"
output:
  pdf_document:
    number_sections: true
    toc: true
    toc_depth: 2
header-includes:
- \usepackage{enumitem}
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "") # omit NAs in Kable
```

\newpage
# Load Packages and Data

```{r message=F, error=F, warning=F}
library(tidyverse)
library(haven)
library(sjlabelled)
library(ggpubr)
library(kableExtra)
library(scales)

# Avoid select clashes
select <- dplyr::select
recode <- dplyr::recode
summarize <- dplyr::summarize
```

\newpage
# Re-Read Project-Level Environmental Variables

```{r}
readRenviron(".Renviron")
```

\newpage
# Import Work, Age, Wave 14 (2018) Flag, and Nursing Home Status

```{r}
# Import "randhrs1992_2018v1.dta"
rand.long <- read_dta(Sys.getenv("HRS_LONG"), 
  col_select = c(hhid, pn, r14agey_e, inw14,
                 r14work, # Currently Working for Pay
                 s14hhidpn)) %>% # Spouse Identifier for Caregiving
  rename(worker = r14work) %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble()

# Import nursing home status from "trk2018tr_r.dta"
tracker <- read_dta(Sys.getenv("HRS_TRACKER_2018_20"),
  col_select = c(hhid, pn, qnurshm))
```

\newpage
# Import Geography Data

```{r}
# Import "HRSXREGION18.dta"
geo <- read_dta(Sys.getenv("HRS_REGION_2018_82"),
    col_select = c(hhid, pn, beale2013_18, region18)) %>%
  rename(rural = beale2013_18) %>%
  mutate(rural = recode(rural, 
    `1` = "Urban",
    `2` = "Urban",
    `3` = "Rural",
    .default = NA_character_)) %>%
  mutate(region = recode(region18, 
    `1` = "northeast", 
    `2` = "northeast",
    `3` = "midwest",
    `4` = "midwest",
    `5` = "south", 
    `6` = "south",
    `8` = "west",
    `9` = "west", 
    .default = NA_character_)) %>%
  mutate(division = recode(region18,
    `1` = "northeast_new_england",
    `2` = "northeast_middle_atlantic",
    `3` = "midwest_east_north",
    `4` = "midwest_west_north",
    `5` = "south_south_atlantic",
    `6` = "south_east_south",
    `7` = "south_west_south",
    `8` = "west_mountain",
    `9` = "west_pacific", .default = NA_character_)) %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble()
```

\newpage
# Import Volunteering Data

```{r}
vol.18 <- read_dta(Sys.getenv("HRS_2018_FAT"),
    col_select = c(hhid, pn, qg086)) %>%
  rename(volunteer = qg086) %>%
  mutate(volunteer = recode(volunteer,
    `1` = 1,
    `5` = 0, 
    .default = NA_real_))
table(vol.18$volunteer)
```

\newpage
# Import Caregiving Data

```{r}
# Import 2018 RAND Fat File (File name: h04f1c.dta)
care18 <- read_dta(Sys.getenv("HRS_2018_FAT"),
                    col_select = c("hhidpn", "hhid", "pn",
                                   
                                   # adl helpers
                                   starts_with("qg033_"),
                                   
                                   # iadl helpers
                                   starts_with("qg055_"),
                                   
                                   # caregiving grandchildren
                                   "qe060", 
                                   
                                   # caregiving parental personal
                                   "qf119", 
                                   
                                   # caregiving parental errands
                                   "qf139")) %>%
  haven::zap_formats() %>%
  sjlabelled::remove_all_labels() %>%
  as_tibble()

# Identify participants who had an ADL or IADL helper
# spouse_helper_sum counts the number of 2s (spouse helper) in qg033 and qg055
# has_spouse_helper is 1 if there is at least one 2s, 0 if none
care18b <- care18 %>%
  mutate(spouse_helper_sum = rowSums(
    ifelse(
      select(., starts_with("qg033") | starts_with("qg055")) == 2, 1, 0
    ), na.rm = TRUE)) %>%
  mutate(has_spouse_helper = ifelse(spouse_helper_sum >= 1, 1, 0))

# Extract participants who have a spousal ADL/IADL caregiver
# Merge their spouse PN
# Then create a dataset with hhid and pn of spouse and an indication of
# whether or not they are a spousal caregiver
spousal_caregivers.18 <- care18b %>%
  filter(has_spouse_helper == 1) %>%
  select(hhid, pn, has_spouse_helper) %>%
  left_join(rand.long %>% select(hhid, pn, s14hhidpn), 
            by = c("hhid", "pn")) %>%
  select(hhidpn = s14hhidpn, caregiver_spousal = has_spouse_helper)

# Merge the spousal_caregivers data back to the dataset
care18c <- care18b %>%
  left_join(spousal_caregivers.18, by = c("hhidpn"))

# Format parental/grandchildren caregivers
# Create caregiver_parental if either personal or errands == 1
care18d <- care18c %>%
  mutate(across(.cols = c(qe060, qf119, qf139), 
    ~recode(., `1` = 1, `5` = 0, `8` = 0, `9` = 0,
            .default = NA_real_))) %>%
  rename(caregiver_grandchildren = qe060,
         caregiver_parental_personal = qf119,
         caregiver_parental_errands = qf139) %>%
  mutate(caregiver_parental =
    ifelse(caregiver_parental_personal == 1 | caregiver_parental_errands == 1,
           1, 0))

# Create general category of caregiving (Note: ignore zeroes)
care18e <- care18d %>%
  select(hhid, pn, cs = caregiver_spousal, cp = caregiver_parental, 
         cg = caregiver_grandchildren) %>%
  mutate(Caregiver_Sum = rowSums(select(., cs:cg), na.rm = T),
         caregiver = ifelse(Caregiver_Sum >= 1, 1, 0)) %>%
  select(hhid, pn, caregiver)
```

\newpage
# Merge Datasets and Create Multiple Productive Activities

```{r}
df <- rand.long %>%
  left_join(tracker, by = c("hhid", "pn")) %>%
  filter(inw14 == 1) %>% # in wave 14
  filter(qnurshm %in% c(5, 6, 7)) %>% # community-dwelling
  filter(r14agey_e >= 65) %>% # age 65+
  left_join(geo, by = c("hhid", "pn")) %>%
  left_join(vol.18, by = c("hhid", "pn")) %>%
  left_join(care18e, by = c("hhid", "pn")) %>%
  mutate(multi = ifelse(volunteer == 1 | caregiver == 1 | worker == 1, 1, 0))
```

## Study Sample Size

```{r}
nrow(df) # 2018 HRS sample of age 65+ community-dwelling individuals
```

\newpage
# Helper Functions

```{r}
# Function for contingency table
get_kab <- function(data, geo, iv) {
  data %>%
    count({{ geo }}, rural, {{ iv }}) %>%
    group_by({{ geo }}, rural) %>%
    filter(!is.na({{ geo }})) %>%
    filter(!is.na(rural)) %>%
    filter(!is.na({{ iv }})) %>%
    mutate(pct = scales::percent(n / sum(n), accuracy = .1)) %>%
    ungroup() %>%
    filter({{ iv }} == 1) %>%
    select({{ geo }}, rural, pct)
}

# Function for two proportion z-test
get_prop <- function(data, geo, iv) {
  data %>%
    count({{ geo }}, rural, {{ iv }}) %>%
    group_by({{ geo }}, rural) %>%
    filter(!is.na({{ geo }})) %>%
    filter(!is.na(rural)) %>%
    filter(!is.na({{ iv }})) %>%
    mutate(sum = sum(n)) %>%
    ungroup() %>%
    filter({{ iv }} == 1) %>%
    pivot_wider(names_from = "rural", values_from = n:sum) %>%
    rowwise() %>%
    mutate(p = prop.test(x = c(n_Rural, n_Urban), 
                         n = c(sum_Rural, sum_Urban))$p.value,
           rural_prop = prop.test(x = c(n_Rural, n_Urban), 
                                  n = c(sum_Rural, sum_Urban))$estimate[1],
           urban_prop = prop.test(x = c(n_Rural, n_Urban), 
                                  n = c(sum_Rural, sum_Urban))$estimate[2]) %>%
    mutate(rural_prop = percent(rural_prop, accuracy = .1),
           urban_prop = percent(urban_prop, accuracy = .1)) %>%
    kbl(booktabs = T, linesep = "", digits = 1) %>%
    kable_styling(position = "center") %>%
    kable_styling(latex_options = c("striped", "hold_position")) 
}
```

\newpage
# Results

## Census Region Statistics

```{r}
r1 <- get_kab(df, region, worker) %>% rename(worker = pct)
r2 <- get_kab(df, region, volunteer) %>% rename(volunteer = pct)
r3 <- get_kab(df, region, caregiver) %>% rename(caregiver = pct)
r4 <- get_kab(df, region, multi) %>% rename(multiple = pct)
r1 %>%
  left_join(r2, by = c("region", "rural")) %>%
  left_join(r3, by = c("region", "rural")) %>%
  left_join(r4, by = c("region", "rural")) %>%
  kbl(booktabs = T, linesep = "", digits = 1) %>%
  kable_styling(position = "center") %>%
  kable_styling(latex_options = c("striped", "hold_position"))  
```

\newpage
## Census Division Statistics

```{r}
d1 <- get_kab(df, division, worker) %>% rename(worker = pct)
d2 <- get_kab(df, division, volunteer) %>% rename(volunteer = pct)
d3 <- get_kab(df, division, caregiver) %>% rename(caregiver = pct)
d4 <- get_kab(df, division, multi) %>% rename(multiple = pct)
d1 %>%
  left_join(d2, by = c("division", "rural")) %>%
  left_join(d3, by = c("division", "rural")) %>%
  left_join(d4, by = c("division", "rural")) %>%
  kbl(booktabs = T, linesep = "", digits = 1) %>%
  kable_styling(position = "center") %>%
  kable_styling(latex_options = c("striped", "hold_position"))  
```

\newpage
## Rural-Urban Comparisons (Two-Proportion Z Tests)

### Region: Worker, Volunteer, Caregiver, and Multiple

```{r}
get_prop(df, region, worker)
get_prop(df, region, volunteer)
get_prop(df, region, caregiver)
get_prop(df, region, multi)
```

\blandscape

\newpage
### Division: Worker and Volunteer

```{r}
get_prop(df, division, worker)
get_prop(df, division, volunteer)
```

\newpage
### Division: Caregiver and Multiple

```{r}
get_prop(df, division, caregiver)
get_prop(df, division, multi)
```

\elandscape